Untitled

### 2024-01-12 manuscript_revised.RMD
getting absolute values RNA
revision work

Data was collected at the sports science lab at the Inland Norway University of Applied Sciences, with all analyses performed on-site, except serum blood analyses, which were performed at Innlandet Hospital Trust. S.E., H.H. and D.H. conceived and designed the study; K.L., S.C.M., D.H. and H.H. performed experiments; K.L. and D.H. interpreted the results; K.L. and D.H. prepared figures; K.L. drafted the manuscript; D.H., H.H. and S.E. edited and revised the manuscript; All authors approved the final version of the manuscript.

@@

Ribosomal biogenesis and content seem to be a prerequisite for skeletal muscle growth, and transcription of ribosomal ribonucleic acid (rRNA) by RNA Polymerase I is considered the rate-limiting step in synthesising new ribosomes [@moss_promotion_1995]. Multiple signalling pathways converge to regulate rRNA transcription, including c-Myc and the mammalian target of rapamycin complex 1 (mTORC1) signal-transduction pathway [@kusnadi_regulation_2015; @von_walden_ribosome_2019; @sebata_control_2018; @west_acute_2016; @mori_c-myc_2021]. The general transcription factor c-Myc increases ribosomal biogenesis directly through transcriptional control of the upstream binding factor (UBF) [@sanij_ubf_2008; @poortinga_c-myc_2011; @mori_c-myc_2021; @west_acute_2016]. Indeed, UBF phosphorylation, which is required for interaction with the ribosomal deoxyribonucleic acid (rDNA) promoter, seems to be increased by high-glucose treatment in a mTORC1-dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells [@mariappan_ribosomal_2011]. Independent of UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures [@zhai_human_2012]. The mTORC1 pathway receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009] and contributes to ribosome biogenesis through the forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription and through the regulation of ribosomal protein translation [@figueiredo_regulation_2019; @von_walden_ribosome_2019]. Moreover, mTORC1 and its downstream target ribosomal protein S6 kinase beta-1 (S6K1) serves as direct mediators of insulin signalling in skeletal muscle [@hillier_physiological_2000]. Previously, increases in rpS6 coincided with increases in rRNA and total RNA in the early phase resistance training in humans [@hammarstrom_ribosome_2022].<!-- FORTSETT HER --> Together these observations indicate a potential role of glucose in positively affecting human skeletal muscle rRNA synthesis, through potentiating transcription or translation of ribosomal RNA and increasing the translational capacity of skeletal muscle cells [@sebata_control_2018; @zhai_human_2012; @hillier_physiological_2000; @hoppe_amp-activated_2009].

@@

The five-session-RT intervention led to increases in total RNA and rRNA per unit muscle weight both in the glucose and the placebo condition, on average by \~20-27% and \~25-57%, respectively (Figure 2). However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.

>>>>

Data was collected at the sports science lab at the Inland Norway University of Applied Sciences, with all analyses performed on-site, except serum blood analyses, which were performed at Innlandet Hospital Trust. D.H., H.H, K.S.M and S.E. conceived and designed the study; D.H., H.H., K.L., K.S.M, and S.C.M. performed experiments; K.L. and D.H. interpreted the results; K.L. and D.H. prepared figures; K.L. drafted the manuscript; D.H., H.H., K.S.M. and S.E. edited and revised the manuscript; All authors approved the final version of the manuscript.

@@

Ribosomal biogenesis and content seem to be a prerequisite for skeletal muscle growth, and transcription of ribosomal ribonucleic acid (rRNA) by RNA Polymerase I is considered the rate-limiting step in synthesising new ribosomes [@moss_promotion_1995]. Multiple signalling pathways converge to regulate rRNA transcription, including c-Myc and the mammalian target of rapamycin complex 1 (mTORC1) signal-transduction pathway [@kusnadi_regulation_2015; @von_walden_ribosome_2019; @sebata_control_2018; @west_acute_2016; @mori_c-myc_2021]. The general transcription factor c-Myc increases ribosomal biogenesis directly through transcriptional control of the upstream binding factor (UBF) [@sanij_ubf_2008; @poortinga_c-myc_2011; @mori_c-myc_2021; @west_acute_2016]. Indeed, UBF phosphorylation, which is required for interaction with the ribosomal deoxyribonucleic acid (rDNA) promoter, seems to be increased by high-glucose treatment in a mTORC1-dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells [@mariappan_ribosomal_2011]. In addition, mTORC1 regulation of ribosomal gene transcription is dependent on ribosomal protein S6 kinase beta-1 (S6K1) [@hannan2003mtor] with its target substrate ribosomal protein S6 (rpS6) coinciding with increases in rRNA and total RNA during early phase RT in humans [@hammarstrom_ribosome_2022]. Independent of UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures [@zhai_human_2012]. The mTORC1 pathway receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009] and contributes to ribosome biogenesis through the forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription and through the regulation of ribosomal protein translation [@figueiredo_regulation_2019; @von_walden_ribosome_2019]. Moreover, mTORC1 and S6K1 serves as direct mediators of insulin signalling in skeletal muscle [@hillier_physiological_2000]. Together these observations indicate a potential role of glucose in positively affecting human skeletal muscle rRNA synthesis, through potentiating transcription or translation of ribosomal RNA and increasing the translational capacity of skeletal muscle cells [@sebata_control_2018; @zhai_human_2012; @hillier_physiological_2000; @hoppe_amp-activated_2009].

@@

The five-session-RT intervention led to increases in total RNA and rRNA per unit muscle weight both in the glucose and the placebo condition, on average by \~20-27% <!--(glucose: 263 ± 50 mg/ng-1, placebo: 210 ± 121 mg/ng-1)--> and \~25-57% <!--(47S; 0.253 ± 1.27 and 0.576 ± 0.677, 18S; 0.336 ± 0.460 and 0.271 ± 0.470, 28S; 0.314 ± 0.504 and 0.311 ± 0.582, 5.8S; 0.388 ± 0.576 and 0.322 ± 0.520, 5S; 0.305 ± 0.608 and 0.292 ± 0.432, arbitrary units for glucose and placebo respectively)--> , respectively (Figure 2). However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.

### 2024-01-12 resources/older manuscript versions/revision_notes.RMD
getting absolute values RNA
revision work

Reviewer 2, comment 2: In the results the data were described in relation to % ou fold of changes. It is important to report the data also in absolute values.
 - Authors reply: Results chapter now updated with absolute values of change from baseline to post in both conditions
   - Glucose revision: Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min: 2.05 ± 0.73 mmol/L), by 31% during RT (Figure 1B, 15 min: 1.75 ± 1.44 mmol/L) and by 32% immediately after RT (Figure 1B, 30 min: 1.62 ± 1.10 mmol/L, all : *p* \< 0.001), with no changes observed in the placebo condition (Figure 1B, 0 min: 0.09 ± 0.3 mmol/L, 15 min: 0.16 ± 0.35 mmol/L, 30 min: 0.18 ± 0.39 mmol/L, *p* \> 0.05). 
   - C-peptide revision: Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min: 796 ± 376.0 pmol/L) and 87% after RT (Figure 1C, 30 min: 793 ± 581.0 pmol/L, both *p* \< 0.001), with no changes observed with the placebo condition (Figure 1C, 0 min: 63.7 ± 71.0 pmol/L, 30 min: 53.9 ± 134.0 pmol/L, both *p* \> 0.05).
   - Absolute values for humac added in a table (tab3), reference by "Table 3 shows the mean change in absolute peak torque values per condition and angular velocity." at the end of the strength index results.
   - Baseline strength values removed from tab 1, since a detailed table of humac measurements has been added (table 3).
   - Mangler å plotte rådata i total RNA, så burde denne kommentaren være besvart.

@@

Reviewer 2, comment 4: Figure 3 shows the result for RPS6. However, only in the discussion section there is some description about this protein (lines 377 - 381). I suggest that this be addressed in the introduction for better understanding of the reader.

@@

>>>>

Reviewer 2, comment 2: In the results the data were described in relation to % ou fold of changes. It is important to report the data also in absolute values.
 - Authors reply: Results chapter now updated with absolute values of change from baseline to post in both conditions
   - Glucose revision: Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min: 2.05 ± 0.73 mmol/L), by 31% during RT (Figure 1B, 15 min: 1.75 ± 1.44 mmol/L) and by 32% immediately after RT (Figure 1B, 30 min: 1.62 ± 1.10 mmol/L, all : *p* \< 0.001), with no changes observed in the placebo condition (Figure 1B, 0 min: 0.09 ± 0.3 mmol/L, 15 min: 0.16 ± 0.35 mmol/L, 30 min: 0.18 ± 0.39 mmol/L, *p* \> 0.05). 
   - C-peptide revision: Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min: 796 ± 376.0 pmol/L) and 87% after RT (Figure 1C, 30 min: 793 ± 581.0 pmol/L, both *p* \< 0.001), with no changes observed with the placebo condition (Figure 1C, 0 min: 63.7 ± 71.0 pmol/L, 30 min: 53.9 ± 134.0 pmol/L, both *p* \> 0.05).
   - Absolute values for humac added in a table (tab3), reference by "Table 3 shows the mean change in absolute peak torque values per condition and angular velocity." at the end of the strength index results.
   - Baseline strength values removed from tab 1, since a detailed table of humac measurements has been added (table 3).
   - Added the following for total RNA and rRNA: (glucose: 263 ± 50 mg/ng-1, placebo: 210 ± 121 mg/ng-1), and; (47S; 0.253 ± 1.27 and 0.576 ± 0.677, 18S; 0.336 ± 0.460 and 0.271 ± 0.470, 28S; 0.314 ± 0.504 and 0.311 ± 0.582, 5.8S; 0.388 ± 0.576 and 0.322 ± 0.520, 5S; 0.305 ± 0.608 and 0.292 ± 0.432, arbitrary units for glucose and placebo respectively)

@@

Reviewer 2, comment 4: Figure 3 shows the result for RPS6. However, only in the discussion section there is some description about this protein (lines 377 - 381). I suggest that this be addressed in the introduction for better understanding of the reader.
 - Authors answer: Added the following;
   - Introduction: "In addition, mTORC1 regulation of ribosomal gene transcription is dependent on ribosomal protein S6 kinase beta-1 (S6K1) [@hannan2003mtor] with its target substrate ribosomal protein S6 (rpS6) coinciding with increases in rRNA and total RNA during early phase RT in humans [@hammarstrom_ribosome_2022]"
   - Discussion:

@@

<!--Also, added Knut Sindre Mølmen to authors, and author contributions (conception of idea, experiments and edition of manuscript)-->
